3D Regression Heat Map Analysis of Population Study Data

### 2015-06-25 draft.tex
added Kais feedback

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
	%%
  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	%%
  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details-on-demand to assess the model quality.
	%%
  \item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details-on-demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
% \noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
% %%
% They extend 2D scatter plots to display detailed information about a regression model.
% %%
% In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
% %%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow, epidemiological data in general and the SHIP in particular.
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
%%
Epidemiological research is performed by experts from different academic disciplines, such as epidemiologists, physicians, statisticians, and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
%%
Derived absolute risks indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
%%
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and are the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
\add{
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model and is suited for comparing models \cite{Akaike}.
%%
A small AIC value indicates a higher quality model.
%%
}
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
%%

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
$Dependent \sim Independent_1 + ... + Independent_n$.
% \begin{equation}
% Dependent \sim Independent_1 + ... + Independent_n
% \label{eq:RegressionNotation}
% \end{equation}
%%
An example of a regression formula would be $KidneyDisorder~\sim~Smoking~+~Obesity$.
The most commonly used regression operators comprise:
%%
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable (e.g. $x\pm y$),
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
\magenta{
Due to the different meaning of operators in regression formulas, feature transformations are not available within this notation.
}
%%
The class of the target feature restricts the regression type.
%%
Different regression types are available, we focus on the following for describing linear relationships:
%%
\paragraph{Linear Regression for Continuous Target.}
%%
The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
%%
The \emph{dependent} variable has to be of \emph{continuous type}.
%%
\add{
%%
Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features.
%%
It yields lower values, when features with little entropy w.r.t. the target feature are included.
%Additional metrics describing linear regression models comprise of the \emph{adjusted} $R^2$, which also adjusts for the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
%%
The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to being not considered at all.
%%
}

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$} \cite{Nagelkerke}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to \emph{derive} hypotheses, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter space analysis using regression models.}
%%
\magenta{
Sedlmair et al.~\cite{Sedlmair} present a conceptual taxonomy of parameter space analysis.
%%
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%%
%Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
%%
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
%%
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
}

\magenta{
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
%%
They provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%%
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%%
%The visualization allows to assess different model complexities.
%%
Existing regression models can be validated and compared using 3D views and 2D slice views.
%%
%We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
%%
M\"{u}hlbacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot matrix view, while we cover more features by abstracting the models.
%%
%We also take categorical targets and regression formula inputs into account.
%%
}

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
%They are therefore suitable for \emph{small} epidemiological data sets.
%%
\magenta{
Maries et al.~\cite{GRACE} propose \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
%%
Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
%%
A multiple view system links spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams (similar to star plots), correlations are plotted using p-values in the other views.
%%
%Dendograms are used to select different brain regions.
Instead of applying several linked views, we focus on one view of abstracted data.
}

@@

The \emph{3D Regression Heat Map} is designed to provide overview visualizations to support hypothesis generation.

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/heat_map}
 \caption{
 (a) Overview visualization using a 2D heat map of the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The $R^2$ metrics extracted from the regression model are mapped to color saturation (a saturated color indicates a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat maps.
 %%
 These represent the slices in our \emph{3D Regression Heat Map}.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Heat_Map}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heat map.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{3D Regression Heat Map} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer the predictive quality of the independent features included in the model.
%%
The $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment (recall Sect.~\ref{sec:RegressionAnalysis}).
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{3D Regression Heat Map} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{3D Regression Heat Map} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (slice) view.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Heat_Map}a).
%%
This encodes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the heat map with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
\emph{Adjusted} $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values have to be normalized in order to map them on color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function color mapping, assigning low $AIC$ features to saturated colors.
%%
}
\design{
%%
To include users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
%To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
%%
}

@@

\paragraph{Selecting and scaling the descriptive regression metric.}
\design{
%%
The feedback made apparent that other features are of interest for analyzing regression models too.
%%
Hence, UI elements for controlling them were introduced.
%%
}
\add{
%%
The descriptive metric shown in the 2D/3D view can be selected using a dropdown menu.
%%
The default selection is $R^2$.
%%
$AIC$ displays model quality.
%%
\emph{Adjusted} $R^2$ values are only available for linear regression.
%%
Logistic regression results are represented via $R^2$ values in this mode.
%%
As they are visually distinguished using color, confusions are avoided.
%%
The transfer function of the color intensity (2D) and opacity (3D) can be adapted using a slider input.
%%
This additionally allows to filter models with desired features, such as only very high $R^2$ values.
}

@@

The purpose of this view is the detailed assessment of the underlying regression models.
%%
By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals \add{f-scores and AIC values}.
%%
\add{
%%
It also contains a scatter plot of the \emph{model residuals}, which shows the difference between the observed data points with the fitted values.
%%
Epidemiologists use such plots to validate models w.r.t. the model assumptions, such as homogeneity, normality, and independence \cite{GraphicModelAssessment}.
}

>>>>

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis generation w.r.t. a target feature using regression models.
	%%
  %\item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	\item Incorporation of prior domain knowledge by using freely adjustable regression formulas.
	%%
  %\item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details-on-demand to assess the model quality.
	\item Metrics selection for analyzing regression models and for details-on-demand representations.
	%%
  %\item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
	\item An open-source web-application that can be used with data of different application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details-on-demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
% \noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
% %%
% They extend 2D scatter plots to display detailed information about a regression model.
% %%
% In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
% %%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow, epidemiological data in general and the SHIP in particular.
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
%%
Epidemiological research is performed by experts from different academic disciplines, such as epidemiologists, physicians, statisticians, and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
%%
Derived absolute risks indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
%%
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and are the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
\add{
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model by estimating its information loss of modeling the underlying data and is suited for comparing models \cite{Akaike}.
%%
A small AIC value indicates a higher quality model.
%%
}
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
%%

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
$Dependent \sim Independent_1 + ... + Independent_n$.
% \begin{equation}
% Dependent \sim Independent_1 + ... + Independent_n
% \label{eq:RegressionNotation}
% \end{equation}
%%
An example of a regression formula would be $KidneyDisorder~\sim~Smoking~+~Obesity$.
The most commonly used regression operators comprise:
%%
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable (e.g. $x\pm y$),
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
\magenta{
Due to the different meaning of operators in regression formulas, feature transformations are not available within this notation.
}
%%
The class of the target feature restricts the regression type.
%%
Different regression types are available, we focus on the following for describing linear relationships:
%%
\paragraph{Linear Regression for Continuous Target.}
%%
The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
%%
The \emph{dependent} variable has to be of \emph{continuous type}.
%%
\add{
%%
Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features.
%%
It yields lower values, when features with little entropy w.r.t. the target feature are included.
%Additional metrics describing linear regression models comprise of the \emph{adjusted} $R^2$, which also adjusts for the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
%%
%The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to being not considered at all.
The $f$-$statistic$ measures the improvement of the model if independent variables were added. 
%%
}

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$} \cite{Nagelkerke}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to \emph{derive} hypotheses, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter space analysis using regression models.}
%%
\magenta{
Sedlmair et al.~\cite{Sedlmair} presented a conceptual taxonomy of parameter space analysis.
%%
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%%
%Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
%%
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
%%
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
}

\magenta{
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
%%
They provided a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%%
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%%
%The visualization allows to assess different model complexities.
%%
Existing regression models can be validated and compared using 3D views and 2D slice views.
%%
%We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
%%
M\"{u}hlbacher et al. focused on a smaller number of features, which can be assessed in more detail, yielding a plot matrix view, while we cover more features by abstracting the models.
%%
%We also take categorical targets and regression formula inputs into account.
%%
}

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
%They are therefore suitable for \emph{small} epidemiological data sets.
%%
\magenta{
Maries et al.~\cite{GRACE} proposed \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
%%
Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
%%
A multiple view system links spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams, correlations are plotted using p-values in the other views.
%%
%Dendograms are used to select different brain regions.
Instead of applying several linked views, we focus on one view of abstracted data.
}

@@

The \emph{3D Regression Heat Map} is designed to provide an overview visualization to support hypothesis generation.

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/heat_map}
 \caption{
 (a) Overview visualization using a 2D heat map of the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The $R^2$ metrics extracted from the regression model are mapped to color saturation (a saturated color indicates a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat maps.
 %%
 These represent the slices in our \emph{3D Regression Heat Map}.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Heat_Map}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heat map.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{3D Regression Heat Map} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer the predictive quality of the independent features included in the model.
%%
The $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment (recall Sect.~\ref{sec:RegressionAnalysis}).
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{3D Regression Heat Map} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{3D Regression Heat Map} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (slice) view.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Heat_Map}a).
%%
This encodes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the heat map with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
\emph{Adjusted} $R^2$ can be represented in the same way, since they are also scaled between $[0, 1]$.
%%
$AIC$ values have to be normalized in order to map them on color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function color mapping, assigning low $AIC$ features to saturated colors.
%%
}
\design{
%%
To include users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
%To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
%%
}

@@

\paragraph{Selecting and scaling the descriptive regression metric.}
\design{
%%
The feedback made apparent that other features are of interest for analyzing regression models too.
%%
Hence, UI elements for controlling them were introduced.
%%
}
\add{
%%
The descriptive metric shown in the 2D/3D view can be selected using a dropdown menu.
%%
The default selection is $R^2$.
%%
$AIC$ displays model quality.
%%
\emph{Adjusted} $R^2$ values are only available for linear regression.
%%
Logistic regression results are represented via $R^2$ values in this mode.
%%
As they are visually distinguished using color, confusions are avoided.
%%
The transfer function of the color intensity (2D) and opacity (3D) can be adapted using a slider input.
%%
This allows to filter models with desired features, such as only very high $R^2$ values.
}

@@

The purpose of this view is the detailed assessment of the underlying regression models.
%%
By hovering over a data entry in the plot, a tooltip displays detailed information about a model's coefficients, associated $p$ values, confidence intervals \add{f-statistics and AIC values}.
%%
\add{
%%
It also contains a scatter plot of the \emph{model residuals}, which shows the difference between the observed data points with the fitted values.
%%
Epidemiologists use such plots to validate models w.r.t. the model assumptions, such as homogeneity, normality, and independence \cite{GraphicModelAssessment}.
}
